### 2016-12-20 paper/chapters/gan.tex
done some stuff
zZzZzzz almost done

%%% when introduced, game theory approach
Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al based on a game theoretic approach to generative modelling.
%which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).

The idea is to have two neural networks playing against each other.
A generator $G(z)$ and a discriminator $D$ where the generator tries to produce data similar to those encountered in a training set. $D$ receives samples from either the training set or the generated data, without knowing where it came from.
Both $G$ and $D$ are usually implemented as neural networks with $G$ being trained to output samples fooling the discriminator and $D$ distinguishing generated samples from training examples.

% use of words: forgery
To understand the inuition behind the approach to GANs, it may be helpful to view(?) analogy of counterfeiters and the police.
Counterfeiters try to produce fake money imitating legit money while the police tries to distinguish fake from real money.
This procedure of generating and distinguishing will go back-and-forth as the counterfeiters will get better at producing authentic looking money and the police is hard-pressed to differentiate between fake and real.
Eventually, this procedure will result in a nash equlibrium where the police or discriminator will have 50\% chance of seperating authentic from real.

Ideally, we want to find this equilibrium for the GAN model as well.

@@

\subsubsection{Minibatch discrimination}
\label{ssub:gan_minibatch_discrimination}

\vspace{10cm}

\subsubsection{Historical Averaging}
\label{ssub:gan_historical_averaging}

>>>>

%%% when introduced, game theory approach
Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modelling.
%which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).

The idea is to have two players playing a non-cooperative game against each other.
One player is called the generator $G$ tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually originating from a dataset.
It is important to note that the generator gets to know whether the discriminator accepted the generated output as \emph{real} or rejected it as \emph{fake}.

Each player plays to maximize its own success, which means the discriminator maximizes the number of samples which got discriminated correctly.
On the other side, $G$ tries to produce as many samples as possibly which got accepted as real by $D$.\\

As the game progresses, both parties will get better at their respective objective and the generated samples will look more and more like the training examples.\\

Using the analogy of counterfeiters and police can help understand the intuition behind GANs better.
Counterfeiters will try to produce fake money looking to fool the police into thinking it is real money.
The police however is able to distinguish badly made fake money from real money easily and as such rejects the fake money with high confidence.
This rejection in turn triggers the counterfeiters to look at real money more closely and try to imitate it even better. Therefore they are able to improve the quality of the faked money. Improvements made to the fake money in this iteration will lower the confidence with which the police was able to reject previous samples.

The game-theoretic solution to the emerging situation where neither counterfeiters as generator nor police as discriminator can improve its own success with respect to the outcome of the game gets called \emph{Nash equilibrium}\cite{game_theory:1994}.

In the case of GANs, this situation means the generator produces samples which can only be rejected with probability $p=0.5$ by the discriminator, meaning it produces as good samples as there are in the dataset.\\\\

To formalize the setting in which the game between $G$ and $D$ takes place, we define the true, however unknown data distribution $p_{data}(x)$ from which samples are available as dataset $X=\{x^{(1)}, x^{(2)}, \dots, x^{(N)}\}$.

In the case of generative modelling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
GANs uses neural network as function approximator for $G$ as well as $D$.
The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples from random noise as $x = G_\xi(z)$,
where $\xi$ denotes the parameters learned by the neural network.\\
Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.

% use of words: forgery
%To understand the inuition behind the approach to GANs, it may be helpful to view(?) analogy of counterfeiters and the police.
%Counterfeiters try to produce fake money imitating legit money while the police tries to distinguish fake from real money.
%This procedure of generating and distinguishing will go back-and-forth as the counterfeiters will get better at producing authentic looking money and the police is hard-pressed to differentiate between fake and real.
%Eventually, this procedure will result in a nash equlibrium where the police or discriminator will have 50\% chance of seperating authentic from real.

%Ideally, we want to find this equilibrium for the GAN model as well.

@@

### 2016-12-20 paper/chapters/introduction.tex
done some stuff
zZzZzzz almost done

Directed generative models are a wide class of models used for generating samples from an unknown distribution.
Lets denote $p_{data}(x)$ as the real-world data distribution which is unknown, however we have access to samples from it: $x \sim p_{data}(x)$.
In order to do so, we define an approximated distribution $p_{model}$ from which we are able to sample.
Under the assumption that $p_{model}$ approximates $p_{data}$ accurately, samples from the true data distribution should have high likelihood under the approximation as well.\\\\

@@

%% on building reasonable representations
The assumption that data can be explained using an underlying, simpler structure than the raw input data is important to handle the high-dimensional nature of most data source.
In order to understand how these statistical models help us to discover the data structure, we will take a look at a few examples of high-dimensional real world data.
\begin{itemize}
  \item \textbf{Images} consisting of millions of pixels can be interpreted as data where each pixel represents an own dimension.
  However, these dimensions are high correlated and structured, for example nearby pixels have with high probability similar values.
  Looking at images of handwritten digits, the data can be reduced to a handful variables which may include the digit number, stroke width, position or size.
  Statistical models help us discover these relations and reduce them to representations in lower dimensional space.
  %\item commonly have thousands of dimensions, but the underlying structure is often far simpler. For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions. There are a few more dimensions, stroke width, cursive, position, size. But still far lower than 784. We will take a look at this example later on in \ref{sec:vae}.
  \item \textbf{Audio data} is omni-present in the real world, however most data sources can be transformed from raw audio samples to spoken words or musical notes without losing too much information for reconstruction.
  \item \textbf{...}
\end{itemize}
% cope with
%To develop systems learning from real world data we use statistical models which are able to infer the underlying structure from
%Real world data is often high-dimensional and noise and transforming this data into an compact representation is therefore an important and also challenging task to perform.
%Generative models are ideal for this task since it has to understand the underlying structure to generate similar data.
%Building this structure is called \textbf{representation learning} and is key to many generative models and how they perform.\\

@@

The assumption that $p(x)$ is dependent on some underlying structure can now be expressed as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors. This stochastic dependency can be represented in a graphical model shown in figure \ref{fig:dgm}.\\

@@

%Inverting $p(x|z)$ yields $p(z|x)$ which is one of the main tasks of some generative models, for example the VAE \ref{sec:vae}.

>>>>

Directed generative models are a wide class of models used for generating samples from an unknown distribution.
Lets denote $p_{data}$ as the unknown real-world data distribution from which samples in the given dataset are drawn.
In order to do so, we define an approximated distribution $p_{model}$ from which we are able to sample.
Under the assumption that $p_{model}$ approximates $p_{data}$ accurately, samples from the true data distribution should have high likelihood under the approximation as well.\\\\

@@

%% on building reasonable representations
%The assumption that data can be explained using an underlying, simpler structure than the raw input data is important to handle the high-dimensional nature of most data source.
Models using deep learning often assume that observed data is stochastically dependent on some hidden structure which we wish to explain.
In order to understand how these statistical models help us to discover the data structure, we will take a look at a few examples of high-dimensional real world data.
\begin{itemize}
  \item \textbf{Images} consisting of millions of pixels can be interpreted as data where each pixel represents an own dimension.
  However, these dimensions are high correlated and structured, for example nearby pixels have with high probability similar values.
  Looking at images of handwritten digits, the data can be reduced to a handful variables which may include the digit number, stroke width, position or size.
  Statistical models help us discover these relations and reduce them to representations in lower dimensional space.
  %\item commonly have thousands of dimensions, but the underlying structure is often far simpler. For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions. There are a few more dimensions, stroke width, cursive, position, size. But still far lower than 784. We will take a look at this example later on in \ref{sec:vae}.
  \item \textbf{Audio data} is omni-present in the real world, however most data sources can be transformed from raw audio samples to spoken words or musical notes without losing too much information for reconstruction.
  %\item \textbf{...}
\end{itemize}
% cope with
%To develop systems learning from real world data we use statistical models which are able to infer the underlying structure from
%Real world data is often high-dimensional and noise and transforming this data into an compact representation is therefore an important and also challenging task to perform.
%Generative models are ideal for this task since it has to understand the underlying structure to generate similar data.
%Building this structure is called \textbf{representation learning} and is key to many generative models and how they perform.\\

@@

The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in figure \ref{fig:dgm}.
%as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors. This stochastic dependency can be represented in a graphical model shown in figure \ref{fig:dgm}.\\

@@

\newpage
%Inverting $p(x|z)$ yields $p(z|x)$ which is one of the main tasks of some generative models, for example the VAE \ref{sec:vae}.

### 2016-12-20 paper/chapters/sigmoid_belief_nets.tex
zZzZzzz almost done

Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985}.
Figure \ref{fig:sbn_arch} shows a bayesian network as a graphical model with each state influenced by its ancestors.
In the case of SBNs however, the states are all binary and the activation function is the sigmoid function. They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
%As described in the overview above, SBNs contain a number of binary states $s$, most commonly divided into many layers.
Each state $s_i$ is a random variables in the bayesian framework which means it can be observed, hidden or unknown representing different states of knowledge. As such it can also be interpreted with a probability distribution.
\begin{equation}
  \label{eq:sbn_node}
p(s_i) = \sigma\bigg(\sum_{j<i}W_{i,j}s_j+b_i\bigg)
\end{equation}
In equation \ref{eq:sbn_node}, $\sigma$ denotes the sigmoid function $\frac{1}{1 + e^{-x}}$ and $W_{i,j}$ the connection weights from state $s_j$ to $s_i$.

To draw samples from this network, it is possible to simply evaluate equation \ref{eq:sbn_node} for each visible node.
However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\
Learning means in this context to adjust the weights and biases of the network to be more likely to generate previously observed input.
Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the input.\\\\

Due to each state being statistically depend on all ancestors, inference means to evaluate all possible configurations of every previous node.
Especially in deep networks, this kind of computation is infeasible.\\

>>>>

Sigmoid Belief Networks (SBN) have been proposed in 1992 by R.M. Neal \cite{neal:1992} representing a specific type of bayesian networks \cite{pearl:1985}.
Figure \ref{fig:sbn_arch} shows a generic belief network as a graphical model with each state influenced by its ancestors.
In the case of SBNs however, all states are binary and the activation function is the sigmoid function. They are one of the earliest neural networks used for generative modelling, predating all other presented models in this article.
%As described in the overview above, SBNs contain a number of binary states $s$, most commonly divided into many layers.
Each state $s_i$ is a random variable in the bayesian framework which means it can be observed, hidden or unknown representing different states of knowledge. As such it can also be interpreted with a probability distribution.
\begin{equation}
  \label{eq:sbn_node}
p(s_i) = \sigma\bigg(\sum_{j<i}W_{i,j}s_j+b_i\bigg)
\end{equation}
In equation \ref{eq:sbn_node}, $\sigma$ denotes the sigmoid function $\frac{1}{1 + e^{-x}}$ and $W_{i,j}$ the connection weights from state $s_j$ to $s_i$.

To draw samples from this network, it is possible to simply evaluate equation \ref{eq:sbn_node} for each visible node.
However prior to that, the SBN first needs to learn how to generate good samples which includes the \emph{learning} and the \emph{inference} problem.\\
Learning means in this context to adjust the weights and biases of the network to be more likely to generate previously observed input.
Meanwhile statistical inference is tasked to infer the states of the hidden nodes based on the input.\\

Due to each state being statistically depend on all ancestors, inference means to evaluate all possible configurations of every previous node.
Especially in deep networks, this kind of computation is infeasible as it requires evaluating exponentially many configurations.\\
